Making (Secure) FETCH Happen: \\ \vspace{2 mm} {\large End-to-End Metadata Security for IMAP}

### 2015-04-29 paper/paper.tex
all of tonight's changes minus acknowledgements

\section{Relevant technologies and related work}

@@

\subsection{Classical email}
Modern email is composed of two distinct protocol stages: sending and transport, and storage and retrieval, shown below in a diagram from \cite{smtp_imap}.

@@

\subsubsection{Messaging Apps}
Some of the most usable and widespread E2E solutions are not for email, but for mobile chat. Apple's iMessage application encrypts messages E2E, although their key management system is unknown and may therefore be invisibly compromised \cite{imessage}. Their messaging system is not only usable: it is a widely-adopted example of a smooth transition to an updated protocol without problematic updates to the user experience. Signal and TextSecure, from Open Whisper Systems, provide a similar experience, but avoid iMessage's public key infrastructure dependency by using instead a modified version of Off-the-Record Messaging (OTR), which can use solutions to the socialist millionaires problem in place of relying on a previously known public key \cite{borisov2004off}. The socialist millionaires problem is a zero-knowledge construction where multiple parties would like to test private values for equality without revealing any information to the other party other than the result of the test. Crucially, Signal and TextSecure enable asynchronous messaging by preemptively generating and sending signed key exchange messages \cite{whisper}. They thereby also provide forward secrecy and deniability in an asynchronous setting. The technology used to implement their solution was implemented in the popular messaging application WhatsApp as well \cite{whatsapp}.

@@

\section{Threat model}

\section{Target user}
\label{targetuser}

\section{Design overview}
To meet all aforementioned requirements, we present \projectnospace, a client-only extension to the IMAP protocol that provides metadata privacy-preserving storage for multiple \project clients. In \projectnospace, emails received at the client are fully encrypted (including mail headers), indexed, and uploaded to a specialized mailbox in the IMAP server. This hides sender and recipient information from the IMAP server. Encryption is based on a passphrase and generated salt, using scrypt as a key derivation function. A special-use ``INDEX'' message is stored in the mailbox as well, and tracks metadata for the client such as the apparent mailbox that a message is in (called a pseudo-mailbox, and visible only to clients, not the server). A special-use ``LOCK'' message is used to allow multiple clients access to the same mailbox, in keeping with IMAP's design principles. Message arrival times are hidden from the server using a combination of scheduling, mixing, fake decoy messages, and techniques drawn from oblivious RAM. \project provides rollback detection for both messages and the index, using version numbers saved client-side and message hashes stored server-side. It is robust to client failure in any state. Details of the design will appear in later sections.

\section{Email Usage Measurements}
To determine various factors such as how often to update and how large a fake message should be, an implementation of \project should measure and dynamically learn these values for actual incoming messages for each particular user. The usage patterns of the author, an American university student, are as follows.

\subsection{Size}

\subsection{Timing}

\section{Architecture and Data Flow}
\label{architecture}

@@

\subsection{Mail From a Secure Source}
\label{mail-from-secure}

@@

\subsection{Mail From an Insecure Source}
\label{mail-from-insecure}

@@

@@

\subsubsection{Scheduling, no mixing}
Could we implement a mix-free solution where we only schedule UP and DOWN operations, hiding actual message additions and deletions among many fake ones? In such a situation, future actions ideally should not reveal information about which messages previously pushed were real and which were fake. Particularly, assuming that deletions of fake messages occur much more frequently than deletions of real messages, deletion operations should reveal neither whether the particular message deleted or any other was real. Additionally, message fetches must similarly not reveal information regarding the veracity of that message or any other. Note that this last restriction immediately disqualifies this solution barring additional mechanisms to mask message fetches by any client of the server that attempts to download all real messages. Further, naive fetches would leak information about which messages are grouped together in a single mailbox. We will see that it is not practical to attempt to find such mechanisms, since the rate of growth required to sufficiently mask deletion operations is infeasible.

@@

With only two clients, if we wish to check our mail more than once every half hour (both of which are conservative estimates), we would use up the entire alloted space in three years. To compensate for space costs, we can introduce mixing operations.

\subsubsection{Mixing with \texorpdfstring{$O(1)$}{O(1)} decoys}

@@

\subsubsection{Mixing with \texorpdfstring{$f(n)$}{f(n)} decoys}
\label{fofn}
With more decoy messages, we could select messages for mixing at random, thereby avoiding issues of distingushing between real and fake messages based on probabilistic distinctions. That is, we could initially seed the server with $f(n)$ messages, and schedule only NONE mixing operations where additional messages appear to neither be pulled nor pushed to the server. Then, when a fake message is randomly selected to appear in the set of messages to be mixed and there is a real message queued for upload to the server, the real message will opportunistically take the place of the fake message in the mix. Likewise, when a real message is selected for deletion, it will be silently replaced with a fake message. The scheduler will then separately decide when to insert an additional fake message in a mix. This insertion is based only on the number of total messages in the server, which the server itself can observe, so no data is leaked there, either.

@@

\subsubsection{ORAM-like protocol}

@@

\section{Attacks and Countermeasures}

Even with all precautions previously described put into place, there are still potential attacks that \project could be vulnerable to.

\subsection{Multiple Client Signup INDEX Rollback}

Given the current SEQUENCE NUMBER mechanism, a server can force a new client to accept a previous version of the system, and use this acceptable to roll back existing clients as well. Here we describe the steps involved in this attack.

@@

\subsubsection{Defense}
When a new client signs up, either manually inspect the contents to ensure that they are in the expected state or display the current sequence number on both an existing client and the new client for a user to compare.

\section{Discussion}
As discussed in Section~\ref{architecture}, every client of an IMAP server must upgrade to \project if any client does so. Therefore, for this technology to spread, \project must be available across platforms on initial release, as a standard modern user expects to be able to check his or her email across devices and email clients. One potentially opportune venue for this is in an enterprise setting. Many businesses require a particular software stack for their employees, often employing IT staff to manage the installation and upkeep on behalf of all workers. Additionally, a business stands to gain much from employing metadata protection. A company involved in manufacturing a product may wish to hide from which vendor it is buying its components before the product's release; disclosure of an email exchange with a particular vendor could compromise this secret. Today, these companies often run their own mail servers to manage such leakages, imposing extraneous time and technology costs on them. With \projectnospace, a company could instead either purchase or build only a mail client, lightening their IT burden.

>>>>

\subsection{Threat Model}

\subsection{Target User}
\label{targetuser}

\section{Relevant Technologies and Related Work}

@@

\subsection{Classical Email}
Modern email is composed of two distinct protocol stages: sending and transport, and storage and retrieval, shown below in a diagram from \cite{smtp_imap}.

@@

\subsubsection{Messaging Apps}
Some of the most usable and widespread E2E solutions are not for email, but for mobile chat. Apple's iMessage application encrypts messages E2E, although their key management system is unknown and may therefore be invisibly compromised \cite{imessage}. Their messaging system is not only usable: it is a widely-adopted example of a smooth transition to an updated protocol without problematic updates to the user experience. Signal and TextSecure, from Open Whisper Systems, provide a similar experience, but avoid iMessage's public key infrastructure dependency by using instead a modified version of Off-the-Record Messaging (OTR), which can use solutions to the socialist millionaires problem in place of relying on a previously known public key \cite{borisov2004off}. The socialist millionaires problem is a zero-knowledge construction where multiple parties would like to test private values for equality without revealing any information to the other party other than the result of the test. Here, the equality test confirms that the two users share a secret. To acquire this shared secret, communicants must exchange or compare information via an out-of-band channel, such as a phone call. So in practice, this verification will only work if users already know each other (otherwise an imposter can carry out a man-in-the-middle attack on the users). Crucially, Signal and TextSecure enable asynchronous messaging by preemptively generating and sending signed key exchange messages \cite{whisper}. They thereby also provide forward secrecy and deniability in an asynchronous setting. The technology used to implement their solution was implemented in the popular messaging application WhatsApp as well \cite{whatsapp}.

@@

\section{System Overview}
To meet all aforementioned requirements, we present \projectnospace, a client-only extension to the IMAP protocol that provides metadata privacy-preserving storage for multiple \project clients. In \projectnospace, emails received at the client are fully encrypted (including mail headers), indexed, and uploaded to a specialized mailbox in the IMAP server. This hides sender and recipient information from the IMAP server. Encryption is based on a passphrase and generated salt, using scrypt as a key derivation function. A special-use ``INDEX'' message is stored in the mailbox as well, and tracks metadata for the client such as the apparent mailbox that a message is in (called a pseudo-mailbox, and visible only to clients, not the server). A special-use ``LOCK'' message is used to allow multiple clients access to the same mailbox, in keeping with IMAP's design principles. Message arrival times are hidden from the server using a combination of scheduling, mixing, fake decoy messages, and techniques drawn from oblivious RAM. \project provides rollback detection for both messages and the index, using version numbers saved client-side and message hashes stored server-side. It is robust to client failure in any state. Details of the design will appear in later sections.

\subsection{Architecture and Data Flow}
\label{architecture}

@@

\subsubsection{Mail From a Secure Source}
\label{mail-from-secure}

@@

\subsubsection{Mail From an Insecure Source}
\label{mail-from-insecure}

@@

\subsection{Email Usage Measurements}
To determine various factors such as how often to update and how large a fake message should be, an implementation of \project should measure and dynamically learn these values for actual incoming messages for each particular user. The usage patterns of the author, an American university student, are as follows.

\subsubsection{Size}

\subsubsection{Timing}

@@

\subsubsection{Scheduling, No Mixing}
Could we implement a mix-free solution where we only schedule UP and DOWN operations, hiding actual message additions and deletions among many fake ones? In such a situation, future actions ideally should not reveal information about which messages previously pushed were real and which were fake. Particularly, assuming that deletions of fake messages occur much more frequently than deletions of real messages, deletion operations should reveal neither whether the particular message deleted or any other was real. Additionally, message fetches must similarly not reveal information regarding the veracity of that message or any other. Note that this last restriction immediately disqualifies this solution barring additional mechanisms to mask message fetches by any client of the server that attempts to download all real messages. Further, naive fetches would leak information about which messages are grouped together in a single mailbox. We will see that it is not practical to attempt to find such mechanisms, since the rate of growth required to sufficiently mask deletion operations is infeasible.

@@

With only two clients, if we wish to check our mail more than once every half hour (both of which are conservative estimates), we would use up the entire allotted space in three years. To compensate for space costs, we can introduce mixing operations.

\subsubsection{Mixing With \texorpdfstring{$O(1)$}{O(1)} Decoys}

@@

\subsubsection{Mixing With \texorpdfstring{$f(n)$}{f(n)} Decoys}
\label{fofn}
With more decoy messages, we could select messages for mixing at random, thereby avoiding issues of distinguishing between real and fake messages based on probabilistic distinctions. That is, we could initially seed the server with $f(n)$ messages, and schedule only NONE mixing operations where additional messages appear to neither be pulled nor pushed to the server. Then, when a fake message is randomly selected to appear in the set of messages to be mixed and there is a real message queued for upload to the server, the real message will opportunistically take the place of the fake message in the mix. Likewise, when a real message is selected for deletion, it will be silently replaced with a fake message. The scheduler will then separately decide when to insert an additional fake message in a mix. This insertion is based only on the number of total messages in the server, which the server itself can observe, so no data is leaked there, either.

@@

\subsubsection{ORAM-like Protocol}

@@

\section{Discussion}

Having motivated, described, and analyzed \projectnospace, we now discuss its limitations, potential plans for rollout, and considerations for future directions.

\subsection{Attacks and Countermeasures}

Even with all precautions previously described put into place, there are still potential attacks that \project could be vulnerable to. Here we give one such attack.

\subsubsection{Multiple Client Signup INDEX Rollback Attack}

Given the current SEQUENCE NUMBER mechanism, a server can force a new client to accept a previous version of the system, and use this acceptable to roll back existing clients as well. The following steps constitute this attack:

@@

\subsubsection{Multiple Client Signup INDEX Rollback Defense}
When a new client signs up, either manually inspect the contents to ensure that they are in the expected state or display the current sequence number on both an existing client and the new client for a user to compare.

\subsection{Dissemination Considerations}

As discussed in Section~\ref{architecture}, every client of an IMAP server must upgrade to \project if any client does so. Therefore, for this technology to spread, \project must be available across platforms on initial release, as a standard modern user expects to be able to check his or her email across devices and email clients. One potentially opportune venue for this is in an enterprise setting. Many businesses require a particular software stack for their employees, often employing IT staff to manage the installation and upkeep on behalf of all workers. Additionally, a business stands to gain much from employing metadata protection. A company involved in manufacturing a product may wish to hide from which vendor it is buying its components before the product's release; disclosure of an email exchange with a particular vendor could compromise this secret. Today, these companies often run their own mail servers to manage such leakages, imposing extraneous time and technology costs on them. With \projectnospace, a company could instead either purchase or build only a mail client, lightening their IT burden.

\subsection{Metadata-Secure Cloud Webmail}
